---
title: "Copepoda"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

An R package to locally serve copepod data from [NOAA NMFS COPEPOD](https://www.st.nmfs.noaa.gov/copepod/) project. This package assists in the download of online datasets and simplifies local management.

Many of the data values are [coded](https://www.st.nmfs.noaa.gov/copepod/codes/), so it is important the user has a working familiarity with the data source.

## Requirements

 + [R v4.1+](https://www.r-project.org/)
 
 + [rlang](https://CRAN.R-project.org/package=rlang)
 
 + [dplyr](https://CRAN.R-project.org/package=dplyr)
 
 + [readr](https://CRAN.R-project.org/package=readr)

 + [sf](https://CRAN.R-project.org/package=sf)
 
## Installation

```
remotes::install_github("BigelowLab/copepoda")
```

## Initial Use

The premise of this package is that data may be stored in a single location but accessed by many users. To achieve this resource-friendly goal and still simplified access for each user, we need to inform the package where the data resides.  We do this by storing the path to the data location in each user's home directory in a hidden text file, "~/.copepoda".  That text file has just one line in it which contains the full path to the shared dataset.  For example, the author's contains `/mnt/ecocast/coredata/noaa/nmfs/copepod` which points to a shared network drive mounted on our linux platform.  

When the package is first loaded (ala `library(copepoda)`) the existence of the file is check, and if missing a warning is issued.

You can create and populate that `~/.copepoda` using a text editors, or you can create using the provided function `set_data_path()`. Here is how the author created his own...

```
library(copepoda)
copepoda::set_data_path("/mnt/ecocast/coredata/noaa/nmfs/copepod")
```

That's it.  If you ever move the data you'll have to modify the contents of this hidden text file.

## Fetching new data
 
To fetch new, or update a local copy, you must know the dataset's ID (aka `copepid`). Below we fetch a dataset from the [Scotian Shelf](https://www.st.nmfs.noaa.gov/copepod/data/ca-01003/index.html) which matches `copepid = ca-01003`.

```{r}
library(copepoda)
ok <- fetch_copepod(copepid = "ca-01003")
```

Now the data directory will contain a subdirectory name `copepod__ca-01003`.  We can list the contents of the [short-form](https://www.st.nmfs.noaa.gov/copepod/data/us-04201/html_src/short-format_intro.html) dataset. 

```{r}
ff <- list_data(copepid = "ca-01003")
head(ff)
```
OK - it only has one file.  Other datasets may have hundreds.

## Reading local data

It's is very simple to read the `short-form` data into a data frame (well, a tibble actually). Make your request by `copepid`... be advised some datasets have many many files and it can take a while to read then all in.

```{r}
x <- read_copepod(list_data(copepid = "ca-01003"), simplify = TRUE)
dplyr::glimpse(x)
```

Alternatively, you can read the data in as a [sf](https://CRAN.R-project.org/package=sf) POINT object.

```{r}
x <- read_copepod(list_data(copepid = "ca-01003"), simplify = TRUE, form = 'sf')
r <- range(x$value_per_vol)
cex <- (x$value_per_vol - r[1])/(r[2]- r[1]) * 3
plot(x['value_per_vol'], axes = TRUE, cex = cex)
```